---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# stdReg2: Regression Standardization for Causal Inference

<!-- badges: start -->
<!-- badges: end -->

Goals: create a unified interface for regression standardization to obtain estimates of causal effects such as the average treatment effect, or relative treatment effect. 

1. Should be easy to use for applied practitioners, i.e., as easy as running glm or coxph. 
2. We want to implement modern, theoretically grounded, doubly-robust estimators, and their associated variance estimators.
3. We want it to be extensible for statistical researchers, i.e., possible to implement new estimators and get other models used within the interface.
4. Robust and clear documentation with lots of examples and explanation of the necessary assumptions. 

## Difference between `stdReg2` and `stdReg`

`stdReg2` is the next generation of `stdReg`. If you are happy using `stdReg`, you can continue using it and nothing will change in the near future. With `stdReg2` we aim to solve similar problems but with nicer output, more available methods, the possibility to include new methods, and mainly to make maintenance and updating easier. 

## Installation

You can install the development version of `stdReg2` from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("sachsmc/stdReg2")
```

## Example

This is a basic example which shows you how to use regression standardization in a logistic regression model to obtain estimates of the causal risk difference and causal risk ratio:

```{r example}
library(stdReg2)

# basic example
# need to correctly specify the outcome model and no unmeasured confounders
# (+ standard causal assumptions)
set.seed(6)
n <- 100
Z <- rnorm(n)
X <- rnorm(n, mean = Z)
Y <- rbinom(n, 1, prob = (1 + exp(X + Z))^(-1))
dd <- data.frame(Z, X, Y)
x <- standardize_glm(
 formula = Y ~ X * Z,
 family = "binomial",
 data = dd,
 values = list(X = 0:1),
 contrasts = c("difference", "ratio"),
 reference = 0
)
x
plot(x)
tidy(x)
```

For more detailed examples, see the vignette "Estimation of causal effects using stdReg2".


## Citation

```{r}
citation("stdReg2")
```

